---
title: "SDM2-Project"
author: "Group 10 - Satya Anudeep Kotaru, Bhavya Teja Inturi, Vineet Sreeram"
date: "2024-05-11"
output: pdf_document
---

## Importing packages

```{r, message = FALSE, warnings = FALSE}
library(tidyverse)
library(data.table)
library(lubridate)
library(plotly)
library(dplyr)
library(stringr)
library(naniar)
library(VIM)
library(FactoMineR)
library(missMDA)
library(ggplot2)
# library(ggfortify)
# library(Amelia)
library(lubridate)
library(forecast)
```

# Importing the Data

```{r}
dt <- fread("household_power_consumption.txt", na.strings = "?")
head(dt)
```

```{r}
sum(is.na(dt))
```

```{r}
dt <- na.omit(dt)
```


### Converting time and date to Time series object
```{r}
dt$datetime <- paste(dt$Date,dt$Time)
dt$datetime <-as.POSIXct(dt$datetime, format="%d/%m/%Y %H:%M:%S")
attr(dt$datetime, "tzone") <- "Europe/Paris"
dt$year <- year(dt$datetime)
dt$week <- week(dt$datetime)
dt$day <- day(dt$datetime)
dt$month <- month(dt$datetime)
dt$minute <- minute(dt$datetime)
```



## time-series 
## time-series plots for sub_metering3
filtered the data for the second week of each month and with the frequency of 52.

```{r}
dt$hour <- hour(dt$datetime)
houseweekly <- filter(dt, week == 2 & hour == 20 & minute == 1)
tsSM3_weekly <- ts(houseweekly$Sub_metering_3, frequency=52, start=c(2007,1))
ggplot2::autoplot(tsSM3_weekly, ts.colour = 'red', xlab = "Time", ylab = "Watt Hours", main = "Sub-meter 3")
```

## time-series plots for sub_metering1
```{r}
tsSM3_weekly <- ts(houseweekly$Sub_metering_1, frequency=52, start=c(2007,1))
ggplot2::autoplot(tsSM3_weekly, ts.colour = 'red', xlab = "Time", ylab = "Watt Hours", main = "Sub-meter 1")
```

## time-series plots for sub_metering2
```{r}
tsSM3_weekly <- ts(houseweekly$Sub_metering_2, frequency=52, start=c(2007,1))
ggplot2::autoplot(tsSM3_weekly,  ts.geom = 'bar', fill = 'blue', xlab = "Time", ylab = "Watt Hours", main = "Sub-meter 2")
```


## time-series plots for sub_metering2 with frequency 20
```{r}
tsSM3_weekly <- ts(houseweekly$Sub_metering_2, frequency=40, start=c(2007,1))
ggplot2::autoplot(tsSM3_weekly,  ts.geom = 'bar', fill = 'blue', xlab = "Time", ylab = "Watt Hours", main = "Sub-meter 2")
```


# Forecasting a time-series
## linear regression
Apply time series linear regression
```{r}
house070809weekly <- filter(dt, year==2008, hour == 20 & minute == 1)
tsSM3_070809weekly <- ts(house070809weekly$Sub_metering_3, frequency=52, start=c(2008,3))
tsSM2_070809weekly <- ts(house070809weekly$Sub_metering_2, frequency=52, start=c(2008,3))
tsSM1_070809weekly <- ts(house070809weekly$Sub_metering_1, frequency=52, start=c(2008,3))

fit3 <- tslm(tsSM3_070809weekly ~ trend + season)
fit2 <- tslm(tsSM2_070809weekly ~ trend + season)
fit1 <- tslm(tsSM1_070809weekly ~ trend + season)
summary(fit3)
```


## Plot the forecast
```{r}
forecastfitSM3c <- forecast(fit3, h=20, level=c(80,90))
forecastfitSM2c <- forecast(fit2, h=20, level=c(80,90))
forecastfitSM1c <- forecast(fit1, h=20, level=c(80,90))
```

```{r}
plot(forecastfitSM3c, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time")
```

```{r}
plot(forecastfitSM2c, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time")
```

```{r}
plot(forecastfitSM1c, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time")
```

# Seasonal Time Series Decomposition
## sub_metering3
```{r}
components070809SM3weekly <- decompose(tsSM3_070809weekly)
plot(components070809SM3weekly)
```

```{r}
summary(components070809SM3weekly)
```

## sub_metering2
```{r}
components070809SM2weekly <- decompose(tsSM2_070809weekly)
plot(components070809SM2weekly)
```

## sub_metering1
```{r}
components070809SM1weekly <- decompose(tsSM1_070809weekly)
plot(components070809SM1weekly)
```

# Applying the exponential prediction
## decomposition
```{r}
tsSM3_070809Adjusted <- tsSM3_070809weekly - components070809SM3weekly$seasonal
tsSM2_070809Adjusted <- tsSM2_070809weekly - components070809SM2weekly$seasonal
tsSM1_070809Adjusted <- tsSM1_070809weekly - components070809SM1weekly$seasonal
autoplot(tsSM3_070809Adjusted)
```

## Check to see the decomposition worked
```{r}
plot(decompose(tsSM3_070809Adjusted))
```

## Fitting the Simple Exponential Smoothing
```{r}
tsSM3_HW070809 <- HoltWinters(tsSM3_070809Adjusted, beta=FALSE, gamma=FALSE)
tsSM2_HW070809 <- HoltWinters(tsSM2_070809Adjusted, beta=FALSE, gamma=FALSE)
tsSM1_HW070809 <- HoltWinters(tsSM1_070809Adjusted, beta=FALSE, gamma=FALSE)
```


## Plot the exponentially smooth 
### Sub_metering3
```{r}
plot(tsSM3_HW070809, ylim = c(0, 25))
```

### Sub_metering2
```{r}
plot(tsSM2_HW070809, ylim = c(0, 25))
```

### Sub_metering3
```{r}
plot(tsSM1_HW070809, ylim = c(0, 25))
```


## Forecast
```{r}
tsSM3_HW070809for <- forecast(tsSM3_HW070809, h=25)
tsSM2_HW070809for <- forecast(tsSM2_HW070809, h=25)
tsSM1_HW070809for <- forecast(tsSM1_HW070809, h=25)
```

### Plot the prediction
#### Sub_metering3
```{r}
plot(tsSM3_HW070809for, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 3")
```

### Plot the prediction
#### Sub_metering2
```{r}
plot(tsSM2_HW070809for, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 2")
```

### Plot the prediction
#### Sub_metering1
```{r}
plot(tsSM1_HW070809for, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 1")
```

```{r}
tsSM3_HW070809forC <- forecast(tsSM3_HW070809, h=25, level=c(10,25))
plot(tsSM3_HW070809forC, ylim = c(0, 20), ylab= "Watt-Hours", xlab="Time - Sub-meter 3", start(2010))
```


#ARIMA

```{r}
dt$Date1 <- as.Date(dt$Date, format = "%d/%m/%Y")
head(dt)
```

```{r}
# Assuming your dataset is named 'data'
# Convert 'Date' and 'Time' columns to datetime format
new_dt <- dt %>%
  mutate(DateTime = as.POSIXct(paste(Date, Time), format = "%d/%m/%Y %H:%M:%S")) %>%
  select(-c(Date, Time))  # Remove original 'Date' and 'Time' columns

# Aggregate data to hourly frequency by taking the mean
hourly_data <- new_dt %>%
  group_by(DateHour = format(DateTime, "%Y-%m-%d %H")) %>%
  summarise(
    Global_active_power = mean(Global_active_power, na.rm = TRUE)
  )
hourly_data <- na.omit(hourly_data)
```


```{r}
hourly_data$Date_Hour <-as.POSIXct(hourly_data$DateHour, format="%Y-%m-%d %H")
hourly_data <- subset(hourly_data, select = -DateHour)
hourly_data <- hourly_data[, c("Date_Hour", "Global_active_power")]
head(hourly_data)
```


```{r}
plot(hourly_data$Date_Hour, hourly_data$Global_active_power, type = "l", xlab = "Date", ylab = "Global Active Power")
```


```{r}
train_end_new <- floor(0.8 * nrow(hourly_data))
train_new <- hourly_data[1:train_end_new, ]
test_new <- hourly_data[(train_end_new + 1):nrow(hourly_data), ]
arima_model_new <- auto.arima(train_new$Global_active_power)
summary(arima_model_new)
```


```{r}
forecast_values_new <- forecast(arima_model_new, h = nrow(test_new))
accuracy(forecast_values_new, test_new$Global_active_power)
```


```{r}
plot(forecast_values_new, main = "ARIMA Forecast", xlab = "Date", ylab = "Global Active Power")

# Add original data for comparison
lines(test_new$Date_Hour, test_new$Global_active_power, col = "blue")
```

# summary
We studied the Household Electric Power Consumption dataset for the time series analysis.
The focus of this study was the sub_metering matters.
The sub_metering category is as follows;
Sub_metering_1 (corresponds to the kitchen, containing mainly a dishwasher, an oven, and a microwave)
Sub_metering_2 (mainly washing-machine, a tumble-drier)
Sub_metering_3 ( water-heater and an air-conditioner)
As it shows, we can check the power consumption of different areas of the houses precisely.


